---
title: "Stat 135 Section 4"
output: html_document
author: GSI - Victor Choi
---

**Exercise 1.**  

#### Bootstrap Confidence Intervals

Given the below sample of 20 data points, and find a Normal, Percentile, and Pivotal confidence interval for log(lambda). 

We believe our sample comes from a Poisson distribution process, but we don't know what the true parameter lambda is.  

```{r}
# Code from previous section

set.seed(1)
# Parametric Bootstrap of Poisson

# Data Set provided
pois_sample <- rpois(20, 3)

# Bootstrap Process - data for each iteration saved in estLogMean
numReps <- 1000
n <- 20

estLogMean <- rep(0, numReps)

for (k in 1:numReps) {
  
  data <- rpois(n, mean(pois_sample))
  estLogMean[k]<-log(mean(data))
}

```

```{r}
# Print 95% Normal, Percentile, and Pivotal confidence intervals for our data

# Add your code below


```



**Exercise 2** 

#### Plotting distributions

For the below 2 data sets, use a boxplot, qqplot, and histogram to understand the difference between the 2 data sets.
Describe briefly the strengths and weaknesses of each plotting method.

```{r}
# Simulate data from two distributions

# T-distribution, 10 degrees of freedom
t_dist = rt(10000, 10, 0)

# Normal distribution
n_dist = rnorm(1000)


```

```{r}
# Add your code below


```


**Exercise 3 (Optional)** 

Repeat Exercise 2 with different numbers of degrees of freedom for the t distribution.





